Towards Explainability of Machine Learning Models in Insurance Pricing

### 2019-10-30 manuscript/manuscript.Rmd
Remove pdf and tex

Arguably, GLMs predate minimum bias procedures by a significant margin. The term 
was coined by @nelder_wedderburn_1972, but generalizations of least squares 
linear regression date back at least to the 1930s. Like minimum bias methods, 
GLMs did not become mainstream in actuarial science for some time. For example, 
the syllabus of basic education does not seem to include any mention of GLMs 
prior to @brown_1988 in the 1990 syllabus for basic education. From there, GLMs 
seem to have received only passing mention until 2006 with the introduction of 
@anderson_2005 to the syllabus.

@@

1. Obtain relevant policy-level data
2. Prepare data for analysis
3. Perform analysis on the data, employing desired method or methods to estimate
needed rate relativites
4. Select final rate relativities based on rate indications
5. Present rates to the reglator, including explanation of the steps followed to 
derive the rates
6. Answer questions from regulators regarding the method employed

The focus of this paper is on steps 5 and 6. In particular, rate regulators are 
concerned with whether rates are inadequate, excessive, or unfairly 
discriminatory. In many states, rate filings that exceed certain thresholds for
magnitude of rate changes or filings that make use of new or sophisticated 
predictive models may be subject to intense regulatory scrutiny. In these cases,
it is necessary to be able to explain the results of the modeling process in a 
way that is understandable without sacrificing statistical rigor.

@@

Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial 
Communications") notes that "...another actuary qualified in the same practice 
area [should be able to] make an objective appraisal of the reasonableness of 
the actuary's work as presented in the actuarial report." [@asop_41] Underlying 
this requirement is an assumption that the hypothetical other actuary qualified 
in the same practice area is adequately familiar with the relevant techniques 
employed. Although the syllabus of basic education is constantly changing, there
has at times been an assumption that all techniques and assumptions that have 
ever been a part of the syllabus of basic education needn't be explained from 
first principles in general actuarial communications, and that an actuary 
practicing in the same field should be able to make an objective appraisal of 
the results from the methods found in the syllabus. [This can be supported by 
reference to the edits to ASOP 38 over time - originally it was designed to be 
about all models, but they revised it to be about only models that incorporate 
specialized knowledge outside of the actuary's area of expertise... do we need 
this citation, though?] This is notable because, beginning with the introduction
of the MAS-I and MAS-II examination in July of 2018, several machine learning 
models were formally included in the syllabus of basic education. These exams 
cover a wide range of topics, such as splines, clustering algorithms, decision 
trees, boosting, and principle components analysis [@cas_syllabus_2018].

Nevertheless, machine learning poses something of a special challenge for ASOP 
41 for several reasons. Machine learning models can be very ad hoc compared to 
traditional statistical models. Because many machine learning models are 
deterministic, they may not admit of standard metrics for model comparison 
(e.g., it's not straightforward to calculate an AIC over a neural network). 
In addition, machine learning methods are often combined into ensembles that 
may not be easily separated and that may, as a collection, cease to resemble 
a single standard version of a model. Complicating matters still further, 
machine learning models can be "black boxes" insofar as the final form of 
response curve cannot be easily predicted and may depend heavily on the 
available data (which may not, in turn, be available to the reviewer).

This last item raises a final interesting issue. GLMs and their ilk are often 
fitted using one of a handful of standard and well-understood approaches (e.g.,
maximum likelihood estimation). However, this is not possible in general with 
machine learning models, as machine learning algorithms often use loss surfaces 
that are very complex such that it may not be feasible to calculate the global 
minimum of the surface. Certainly, closed form representations of the loss 
surfaces are not generally available. For this reason, the training phase of a 
machine learning model is, in many ways, just as important to one's 
understanding as the model form and the data on which the model is fitted. 
Because the final model result is inseparable from these three components 
(training method, model form, and data), it is not generally adequate to just
know the method employed to make an objective appraisal of the reasonableness 
of the model. More information is necessary.

Of course, these comments only apply to the actuarial profession. Outside of 
the actuarial profession, communication of results may be more challenging. A 
2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
the National Association of Insurance Commissioners found that the plurality of
responding regulators identified "Filing complexity and/or a lack of resources 
or expertise" as a key challenge that impedes their ability to review GLMs or 
other predictive models [@naic_summer_2017]. Given that machine learning 
algorithms are generally regarded as more complex than GLMs, this implies that
the challenge of communicating machine learning model results is significant.

@@

* Determine the extent the model causes premium disruption for individual 
policyholders, and how the insurer will
explain the disruption to individual consumers that inquire about it.
* Determine that individual input characteristics to a predictive model are 
related to the expected loss or expense
differences in risk. Each input characteristic should have an intuitive or 
demonstrable actual relationship to expected loss or expense.
* Determine that individual outputs from a predictive model and their associated
selected relativities are not unfairly discriminatory.

@@

We motivate our discussion by considering several aspects of interpretability. 
As we proceed through the points below, we aim to arrive at a more scoped and 
relevant definition of what it means to attain interpretability in a pricing
model. In the remainder of this section, we clarify a couple concepts regarding 
interpretable *classes* of models and and the computational transparency of ML
models, outline findings on the social aspects of ML explanations most relevant
to our use case, then discuss a potential framework for implementing ML 
interpretability in practice.

@@

In the actuarial science literature, the GLM is probably the most oft-cited
example of an easily interpretable model. Given a set of inputs, we can easily
reason about what the output of the model is. As an illustrative example, 
consider a claim severity model with driver age, sex, and vehicle age as 
features; assuming a log link function and letting $Y$ denotes the response, we
have

@@

Here, we can tell, for example, what the model would predict for the expected 
severity if we were to increase age by a certain amount, *all else being equal*,
because the relationship between the feature and the response is simply 
multiplication by the coefficient $\beta_1$ and applying the inverse link
function.

@@

Now, it is worth pointing out that, when declaring that GLMs or decision trees 
are interpretable models, we are implicitly assuming that we are considering 
only a handful  of features. In fact, the ease at which we can reason about a
model declines as the number of features, transformations of them, and
interactions increase, as in the following (somewhat pathological) example:

@@

Proposed by @hiltonConversationalProcesses1990 and interpreted by
@millerExplanationArtificial2017 in the context of ML, model explanations
are conversations or social interactions. One consequence of this identification
is that explanations need to be *relevant* to the *audience*. In developing,
filing, and operationalizing a pricing model, one needs to accommodate a variety
of stakeholders, each of whom has a different set of questions, assumptions,
and technical capacity. First, there are internal stakeholders at the company,
which includes management and underwriters. While some of the individuals in
this audience may be technical, they are likely less familiar with predictive
modeling techniques than the actuaries and data scientists who build the models.
Next, we have the regulators, who may have limited resources to review the
models, and will focus on a specific list of questions motivated by statute
and public policy. Finally, we have the public, who has an interest (perhaps
more so than the other parties) because their lives are directly affected by
the premiums charged.

It is interesting to note that, the modelers, who are most familiar with the
models, tend to be same people designing and communicating the explanations.
This poses a challenge that @millerExplainableAI2017 calls "inmates running
the asylum", where the modelers design explanations for *themselves* rather
than the intended audience. For example, they may be interested in technical
questions such as extrapolation behavior and shape the explanations accordingly,
which may be lost on the prospective policyholder.

@@

These questions should relate to existing guidelines, such as those described in
@naic_white_paper and outlined in Section \ref{need}, standards of practice, and
regulation, and in fact should not be specific only to ML models. By
standardizing on a set of questions, we reduce the burden of both companies and
regulators; this is especially important for the latter, who are already
resource constrained facing increasing variety of models being filed. This
format should also be familiar to actuaries who are accustomed to adhering to
specific guidelines in, for example, ASOPs. Like the ASOPs, We envision that
these questions and guidelines will be continually updated to reflect feedback
obtained and advances in research.

@@

1. What are the most important predictors in the model?
2. How does the predicted loss cost change, on average, as we change an input?
3. For a particular policyholder, how does each characteristic contribute to
the loss cost prediction?

@@

For linear models and their generalizations, and some ML models, measures of
variable importance can be obtained from the fitted model structure. In the
case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas
for random forests, one can use out-of-bag errors [@breimanRandomForests2001].
For more complex models, such as the neural network in our case study, we need
to devise another approach.

@@

More formally, let $y$ denote the vector of responses, $X$ denote the matrix of
predictor variables, $\widehat{f}$ denote the fitted model, and
$L = \mathcal{L}(\widehat{f}(X), y)$, where $\widehat{f}$ applies to $X$
rowwise, denote the value of the loss function, which is mean squared error
in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor
matrix where the $j$th variable has been permutated, then we can compute the
loss with the permuted dataset as
$L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. With this, we define
the  variable importance $VI^{j}$ as $L^{-j} - L$.

In Figure \@ref(fig:fi), we show a plot of variable importances. In our
particular example, we see that the "make" variable contributes most to the
accuracy of the model with "sex" contributing the least. This provides a way
for the audience to quickly glance at the most relevant variables, and ask
further question as necessary. For example, if a variable that can potentially
be a proxy for a discriminatory variable appears as very important, more
investigation would be warranted.

@@

In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable.
We see that the average predicted loss cost decreases with vehicle age until
the latter is around 18. Note that the rug on the $x$-axis tell us that the
data is quite thin for vehicle age greater than 18, so the apparent upward
trend to the right is driven by just a few data points. With this in mind,
we can conclude that as cars get older, we expect the loss cost predictions
generated by our model to decrease.

@@

(TODO)

@@

In this appendix, we describe the ML model and the data used to train it. Note 
that, for our paper, the ultimate goal of the modeling procedure is to develop 
something that can produce predictions. As a result, we do not follow standard
practices for supervised learning model development. However, for the sake of 
completeness and reproducibility, we include an overview of the process here. 
Implementation is done using the R [@rlang] interface to TensorFlow 
[@tensorflow2015-whitepaper]. The model explanation visualizations utilize the 
implemention by @biecekPMVEE, and the code to reproduce them are available on
GitHub[^1].

@@

Table \@ref(tab:variables) shows the input variables to our model and their 
associated transformations. For "make" and "region", we map each level to a 
point in $\mathbb{R}^2$ through embedding layers. The model is a feedforward
neural network with two hidden layers with 64 units each. The activations for
the hidden layers are ReLU while for the output layer it is softplus. Exposures
for each record are used as sample weights during training. We fit the model
via ADAM with an initial learning rate of 0.1, a mini-batch size of 10,000,
and trigger early stopping when the mean squared error on the asssessment set 
does not improve for five epochs.

>>>>

Arguably, GLMs predate minimum bias procedures by a significant margin. The term 
was coined by @nelder_wedderburn_1972, but generalizations of least squares 
linear regression date back at least to the 1930s. Like minimum bias methods, 
GLMs did not become mainstream in actuarial science for some time. For example, 
the syllabus of basic education of the Casualty Actuarial Society does not seem to include any mention of GLMs 
prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
seem to have received only passing mention until 2006 with the introduction of 
@anderson_2005 to the syllabus.

@@

1. Obtain relevant policy-level data
2. Prepare data for analysis
3. Perform analysis on the data, employing desired method(s) to estimate
needed rates
4. Select final rates based on rate indications
5. Present rates to the reglator, including explanation of the steps followed to 
derive the rates
6. Answer questions from regulators regarding the method employed

The focus of this paper is on steps 5 and 6. In particular, rate regulators are 
concerned with whether rates are inadequate, excessive, or unfairly 
discriminatory. In many states, rate filings that exceed certain thresholds for
magnitude of rate changes or filings that make use of new or sophisticated 
predictive models may be subject to particular regulatory scrutiny. In these cases,
it is necessary to be able to explain the results of the modeling process in a 
way that is understandable without sacrificing statistical rigor.

@@

Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial 
Communications") notes that "...another actuary qualified in the same practice 
area [should be able to] make an objective appraisal of the reasonableness of 
the actuary's work as presented in the actuarial report." [@asop_41] Underlying 
this requirement is an assumption that the hypothetical other actuary qualified 
in the same practice area is adequately familiar with the relevant techniques 
employed. Although the syllabus of basic education is constantly changing, there
has at times been an assumption that all techniques and assumptions that have 
ever been a part of the syllabus of basic education needn't be explained from 
first principles in general actuarial communications, and that an actuary 
practicing in the same field should be able to make an objective appraisal of 
the results from the methods found in the syllabus.
This is notable because, beginning with the introduction
of the MAS-I and MAS-II examination in July of 2018, several machine learning 
models were formally included in the syllabus of basic education. These exams 
cover a wide range of topics, such as splines, clustering algorithms, decision 
trees, boosting, and principle components analysis [@cas_syllabus_2018].

Nevertheless, machine learning poses something of a special challenge for ASOP 
41 for several reasons:

1. Machine learning models can be very ad hoc compared to 
traditional statistical models.
2. Because many machine learning models do not assume an underlying probability
distribution or stochastic process, they may not admit of standard metrics for model comparison 
(e.g., it's not straightforward to calculate an AIC over a neural network). 
3. Machine learning methods are often combined into ensembles that 
may not be easily separated and that may, as a collection, cease to resemble 
a single standard version of a model.
4. Machine learning models can be "black boxes" insofar as the final form of 
response curve cannot be easily predicted and may depend heavily on the 
available data (which may not, in turn, be available to the reviewer).

This last item raises a final interesting issue. GLMs and their ilk are often 
fitted using one of a handful of standard and well-understood approaches (e.g.,
maximum likelihood estimation). However, this is not possible in general with 
machine learning models, as machine learning algorithms often use loss surfaces 
that are very complex such that it may not be feasible to calculate the global 
minimum of the surface. Certainly, closed form representations of the loss 
surfaces are not generally available. For this reason, the training phase of a 
machine learning model is, in many ways, just as important to one's 
understanding as the model form and the data on which the model is fitted. 
Because the final model result is inseparable from these three components 
(training method, model form, and data), it is not generally adequate to just
know the method employed to make an objective appraisal of the reasonableness 
of the result. More information is necessary.

Of course, these comments only apply within the actuarial profession. Outside of 
the actuarial profession, communication of results may be more challenging. A 
2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
the National Association of Insurance Commissioners found that the plurality of
responding regulators identified "Filing complexity and/or a lack of resources 
or expertise" as a key challenge that impedes their ability to review GLMs or 
other predictive models [@naic_summer_2017]. Given that machine learning 
algorithms are generally regarded as more complex than GLMs, this implies that
the challenge of communicating machine learning model results is significant.

@@

* Determine the extent to which the model causes premium disruption for individual 
policyholders and how the insurer will
explain the disruption to individual consumers that inquire about it.
* Determine that individual input characteristics to a predictive model are 
related to the expected loss or expense
differences in risk. Each input characteristic should have an intuitive or 
demonstrable actual relationship to expected loss or expense.
* Determine that individual outputs from a predictive model and their associated
selected relativities are not unfairly discriminatory.

@@

We motivate our discussion by considering several aspects of interpretability. 
As we proceed through the points below, we aim to arrive at a more scoped and 
relevant definition of what it means for a pricing
model to be interpretable. In the remainder of this section, we clarify a couple concepts regarding 
interpretable *classes* of models and the computational transparency of ML
models, outline frameworks for understanding the communication goals of interpretability,
then discuss a potential framework for implementing ML 
interpretability in practice.

@@

In the actuarial science literature, the GLM is probably the most oft-cited
example of an easily interpretable model. Given a set of inputs, we can easily
reason about what the output of the model is. As an illustrative example, 
consider a claim severity model with driver age, sex, and vehicle age as 
predictors; assuming a log link function and letting $Y$ denote the response, we
have

@@

Here, we can tell, for example, what the model would predict for the expected 
severity if we were to increase age by a certain amount, *all else being equal*,
because the relationship between the predictor and the response is simply 
multiplication by the coefficient $\beta_1$ and applying the inverse link
function.

@@

Now, it is worth pointing out that, when declaring that GLMs or decision trees 
are interpretable models, we are implicitly assuming that we are considering 
only a handful  of predictors. In fact, the ease with which we can reason about a
model declines as the number of predictors, transformations of them, and
interactions increase, as in the following (somewhat pathological) example:

@@

@hiltonConversationalProcesses1990 proposed a framework, later interpreted by
@millerExplanationArtificial2017 in the context of ML, for understanding model explanations
as *conversations* or *social interactions.* One consequence of this identification
is that explanations need to be *relevant* to the *audience*. This framework is
consistent with ASOP 41, which formulates a similar requirement in terms of an
*intended user* of the actuarial communication. In developing,
filing, and operationalizing a pricing model, one needs to accommodate a variety
of stakeholders, each of whom has a different set of questions, assumptions,
and technical capacity. First, there are internal stakeholders at the company,
which includes management and underwriters. While some of the individuals in
this audience may be technical, they are likely less familiar with predictive
modeling techniques than the actuaries and data scientists who build the models.
Next, we have the regulators, who may have limited resources to review the
models, and will focus on a specific list of questions motivated by statute
and public policy. Finally, we have potential policyholders, who have an interest (perhaps
more so than the other parties) as they are responsible for paying the premiums.

It is interesting to note that the modelers, who are most familiar with the
models, tend to be same people designing and communicating the explanations.
This poses a challenge that @millerExplainableAI2017 calls "inmates running
the asylum", where the modelers design explanations for *themselves* rather
than the intended audience. For example, they may be interested in technical
questions, such as extrapolation behavior, and shape the explanations accordingly,
which may be irrelevant to a prospective policyholder.

@@

It should be noted that no ideal set of questions exists that would encompass
all potential models. Rather, the actuary must consider what aspects of the
model would raise questions from the perspective of the model's intended users.
We propose that relevant stakeholders, by providing example questions and
answers, would inherently provide guidance by which actuaries can reasonably
anticipate the kinds of specific questions most important to those stakeholders
and address them proactively.

These questions should relate to existing guidelines, such as those described in
@naic_white_paper and outlined in Section \ref{need}, standards of practice, and
regulation, and in fact should not be specific only to ML models. By
conceptualizing a set of questions, we reduce the burden of both companies and
regulators; this is especially important for the latter, who are already
resource constrained facing increasing variety of models being filed. This
format should also be familiar to actuaries who are accustomed to adhering to
specific guidelines in, for example, ASOPs. Like the ASOPs, We envision that
these questions and guidelines will be continually updated to reflect feedback
obtained and advances in research.

@@

1. What are the most important predictors in the model? Put another way, to
what extent do the predictors improve the accuracy of the model?
2. How does the predicted loss cost change, on average, as we change an input?
3. For a particular policyholder, how does each characteristic contribute to
the loss cost prediction?

@@

For linear models and their generalizations, and some ML models, measures of
variable importance can be obtained from the fitted model structure. In the
case of GLMs, one might observe the magnitudes of the $t$-statistics, step-wise
reduction in information criteria, or out-of-sample performance, whereas
for random forests, one might use out-of-bag errors [@breimanRandomForests2001].
For more complex models, such as the neural network in our case study, we need
to devise another approach.

@@

More formally, let $y$ denote the vector of responses, $X$ denote the matrix of
predictor variables, $\widehat{f}$ denote the fitted model, and
$L = \mathcal{L}(\widehat{f}(X), y)$, where $\widehat{f}$ applies to $X$
rowwise, denote the value of the loss function, which is mean squared error
in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor
matrix where the $j$th variable has been permuted, then we can compute the
loss with the permuted dataset as
$L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. With this, we define
the  variable importance $VI^{j}$ as $L^{-j} - L$.

In Figure \@ref(fig:fi), we show a plot of variable importances. In our
particular example, we see that the "make" variable contributes most to the
accuracy of the model with "sex" contributing the least. This provides a way
for the audience to quickly glance at the most relevant variables, and ask
further question as necessary.

@@

Note that this proposed plot can be calculated over the dataset or over a
hypothetical dataset in which the value of $x_i^{j}$ is set equal to $z$ for
each record.

In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable.
We see that the average predicted loss cost decreases with vehicle age until
the latter is around 18. Note that the rug on the $x$-axis tell us that the
data is quite thin for vehicle age greater than 18, so the apparent upward
trend to the right is driven by just a few data points.

This information allows the modeler and stakeholders to consider whether it
is reasonable for the anticipated loss cost to follow this shape.

@@

Actuarial standards of practice, most notably ASOP 41, places responsibility on
the actuary to clearly communicate actuarial work products, including insurance
pricing models. These responsibilities create special challenges for
communicating machine learning models, which are often seen "black boxes" due
in part to their complexity, nonlinearity, flexible construction, and ad hoc
nature.

In this paper, we discussed particular questions of model validation that are
of key importance in communicating a model that may present particular
difficulty for machine learning models compared to GLMs or traditional pricing
models. Specifically,

* How does the model impact individual insurance consumers?
* How are the predictor variables related to expected losses?
* Are the variables unfairly discriminatory?

We contextualized these questions in terms of different frameworks for defining
interpretability. We conceptualize interpretability in terms of the ability of
a model (or modeler) to answer a set of idealized questions that would be
refined. We then offered potential (families of) model-agnostic techniques for
providing answers to these questions.

Much work remains to be done in terms of defining the role of machine learning
algorithms in actuarial practice. Lack of interpretability has been a key
barrier preventing wider adoption and exploration of these techniques. The
methods proposed in this paper could therefore represent important strides in
nlocking the potential of machine learning within the insurance industry.

@@

In this appendix, we describe the ML model and the data used to train it. Note 
that, for our paper, the ultimate goal of the modeling procedure is to develop 
something that can produce predictions. As a result, we do not follow standard
practices for tuning and validation. However, for the sake of 
completeness and reproducibility, we include an overview of the process here. 
Implementation is done using the R [@rlang] interface to TensorFlow 
[@tensorflow2015-whitepaper]. The model explanation visualizations utilize the 
implemention by @biecekPMVEE, and the code to reproduce them are available on
GitHub[^1].

@@

Table \@ref(tab:variables) shows the input variables to our model and their 
associated transformations. For "make" and "region", we map each level to a 
point in $\mathbb{R}^2$ through embedding layers. The model predicts expected
loss cost for all perils combined. The architecture is a feedforward
neural network with two hidden layers with 64 units each. The activations for
the hidden layers are ReLU while for the output layer it is softplus. Exposures
for each record are used as sample weights during training. We fit the model
via ADAM with an initial learning rate of 0.1, a mini-batch size of 10,000,
and trigger early stopping when the mean squared error on the asssessment set 
does not improve for five epochs.
